---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Tanguy Levent"
date: "26 février 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
header-includes: \usepackage{color}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

I download the ToothGrowth data in the R datasets package so that analyze the effect of the vitamin C on the Tooth growth in Guinea pigs.
First I'm going to provide basic summarises to understand clearly the variables and observations of the dataset.
Then I perform some basic exploratory data analyses to finally use confidence intervals and/or hypothesis tests.

Before to go forward, please to take a look on the [\textcolor{blue}{ToothGrowth R package description}](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html)


## 1. Load the ToothGrowth data  

I load my dataset and store it in a variable called *"mdata"*.  

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(RColorBrewer)

data("ToothGrowth")
mdata <- ToothGrowth
sum(is.na(mdata))
```

**Note that there are no missing values in our dataset.**  

## 2. Provide a basic summary of the data  

```{r }
head(mdata)
```   

The dataset is composed of the 3 variables : *len* , *supp* and *dose*  
  
  -----

```{r}
str(mdata)
```  

The dataset contains 60 observations. The variable *supp* which represents the delivery methods, 
have 2 levels inside : "OJ" = "Orange Juice" and "CV" = "Ascorbic Acid".


## 3. Perform some basic exploratory data analyses  

We have understood that the measures taken in the dataset represents tooth growth compared to the dose of vitamin C and
the delivery methods. To understand the split by subject/observation I compute a table between *dose* and *supp* :  

```{r}
table(mdata$dose, mdata$supp)
```  

In this way, we can observed that the dataset splits 30 pigs by delivery methods and as we have 3 dose levels for each observation, there are 10 subjects by dose levels.  

*For example we have 10 pigs which had 0.5 mg/day of vitamin C given by Orange Juice method*  


Now we're going to go more deeper into the analyzes. It will be interesting to see if the dose levels of vitamin C 
have an impact on the tooth growth. For that I realize with ggplot function a boxplot to vizualise the study (code in Appendix):  

  
```{r,echo=FALSE}

ggplot(mdata, aes(factor(dose), len)) + 
        geom_boxplot(aes(fill = factor(dose))) +
        scale_fill_brewer(palette="Pastel1") + 
        xlab("Dose levels of Vitamin (mg/day)") + 
        ylab("Length of Tooth") + 
        ggtitle("Effect of dose levels of Vitamic C on the Tooth Growth") + 
        guides(fill=guide_legend(title="Dose Levels"))+
        theme(plot.title = element_text(hjust = 0.5))

```  

\pagebreak
Write the mean values by dose levels :  


```{r,echo=FALSE}
Global_Dose <- mdata %>% group_by(dose) %>% summarise(len = mean(len))
Global_Dose
```  
Quickly, thank to the boxplot and the table above, we can conclude that indeed as more the pigs have a daily dose of vitamin C, as more their tooth growth is important.  

Now we want to see if the delivery methods of vitamin C as well have an impact on the tooth growth of the pigs. I take the same proceed to realize it :  
  
```{r,echo=FALSE}

ggplot(mdata, aes(factor(dose), len)) + 
        geom_boxplot(aes(fill = factor(supp))) + 
        facet_grid( ~ supp) +  
        scale_fill_brewer(palette="Pastel1") + 
        xlab("Dose levels of Vitamin C (mg/day)") + 
        ylab("Length of Tooth")  + 
        ggtitle("Comparaison effect of dose levels of Vitamin C on the Tooth Growth by Delivery Methods") +
        guides(fill=guide_legend(title="Delivery Methods")) + 
        theme(plot.title = element_text(size = 09 , hjust = 0.5))

```  

```{r,echo=FALSE}
splitbySupp <- mdata %>% group_by(supp,dose) %>% summarise(len = mean(len))
Global_Supp_Mean <- splitbySupp %>% group_by(supp) %>% summarise(len = mean(len))
Global_OJ <- splitbySupp %>% filter(supp=="OJ")
Global_VC <- splitbySupp %>% filter(supp=="VC")



```  
   
**- Means in terms of delivery methods**  

```{r,echo=FALSE}
Global_Supp_Mean
```  

We can conclude that effectively the delivery methods influence the lenght of tooth at each degree of dosage. With the table
above we understand that feed pigs of vitamin C with Orange Juice will produce better tooth growth than with Ascorbic Acid.

\pagebreak  

**- Means of Orange Juice by dose**  

```{r,echo=FALSE}
Global_OJ

```  

**- Means of Ascorbic Acid by dose**  

```{r,echo=FALSE}
Global_VC
```  

However, note that at 2 mg/day, there is no difference in the tooth growth between OJ or VC absorption (26.06 VS 26.14).  

## 4. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose


