---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Tanguy Levent"
date: "26 février 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
header-includes: \usepackage{color}
---

## Synopsis

I download the ToothGrowth data in the R datasets package so that analyze the effect of the vitamin C on the Tooth growth in Guinea pigs.
First I'm going to provide basic summarises to understand clearly the variables and observations of the dataset.
Then I perform some basic exploratory data analyses to finally use confidence intervals and/or hypothesis tests.

Before to go forward, please to take a look on the [\textcolor{blue}{ToothGrowth R package description}](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html)


## 1. Load the ToothGrowth data  

I load my dataset and store it in a variable called *"mdata"*.  

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(RColorBrewer)

data("ToothGrowth")
mdata <- ToothGrowth
sum(is.na(mdata))
```

**Note that there are no missing values in our dataset.**  

## 2. Provide a basic summary of the data  

```{r }
head(mdata)
```   

The dataset is composed of the 3 variables : *len* , *supp* and *dose*  
  
  -----

```{r}
str(mdata)
```  

The dataset contains 60 observations. The variable *supp* which represents the delivery methods, 
have 2 levels inside : "OJ" = "Orange Juice" and "CV" = "Ascorbic Acid".


## 3. Perform some basic exploratory data analyses  

We have understood that the measures taken in the dataset represents tooth growth compared to the dose of vitamin C and
the delivery methods. To understand the split by subject/observation I compute a table between *dose* and *supp* :  

```{r}
table(mdata$dose, mdata$supp)
```  

In this way, we can observed that the dataset splits 30 pigs by delivery methods and as we have 3 dose levels for each observation, there are 10 subjects by dose levels.  

*For example we have 10 pigs which had 0.5 mg/day of vitamin C given by Orange Juice method*  


Now we're going to go more deeper into the analyzes. It will be interesting to see if the dose levels of vitamin C 
have an impact on the tooth growth. For that I realize with ggplot function a boxplot to vizualise the study (code in Appendix):  

  
```{r,echo=FALSE}

ggplot(mdata, aes(factor(dose), len)) + 
        geom_boxplot(aes(fill = factor(dose))) +
        scale_fill_brewer(palette="Pastel1") + 
        xlab("Dose levels of Vitamin (mg/day)") + 
        ylab("Length of Tooth") + 
        ggtitle("Effect of dose levels of Vitamic C on the Tooth Growth") + 
        guides(fill=guide_legend(title="Dose Levels"))+
        theme(plot.title = element_text(hjust = 0.5))

```  

\pagebreak
Write the mean values by dose levels :  


```{r,echo=FALSE}
Global_Dose <- mdata %>% group_by(dose) %>% summarise(len = mean(len))
Global_Dose
```  
Quickly, thank to the boxplot and the table above, we can conclude that indeed as more the pigs have a daily dose of vitamin C, as more their tooth growth is important.  

Now we want to see if the delivery methods of vitamin C as well have an impact on the tooth growth of the pigs. I take the same proceed to realize it :  
  
```{r,echo=FALSE}

ggplot(mdata, aes(factor(dose), len)) + 
        geom_boxplot(aes(fill = factor(supp))) + 
        facet_grid( ~ supp) +  
        scale_fill_brewer(palette="Pastel1") + 
        xlab("Dose levels of Vitamin C (mg/day)") + 
        ylab("Length of Tooth")  + 
        ggtitle("Comparaison effect of dose levels of Vitamin C on the Tooth Growth by Delivery Methods") +
        guides(fill=guide_legend(title="Delivery Methods")) + 
        theme(plot.title = element_text(size = 09 , hjust = 0.5))

```  

```{r,echo=FALSE}
splitbySupp <- mdata %>% group_by(supp,dose) %>% summarise(len = mean(len))
Global_Supp_Mean <- splitbySupp %>% group_by(supp) %>% summarise(len = mean(len))
Global_OJ <- splitbySupp %>% filter(supp=="OJ")
Global_VC <- splitbySupp %>% filter(supp=="VC")



```  
   
**- Means in terms of delivery methods**  

```{r,echo=FALSE}
Global_Supp_Mean
```  

We can conclude that effectively the delivery methods influence the lenght of tooth at each degree of dosage. With the table
above we understand that feed pigs of vitamin C with Orange Juice will produce better tooth growth than with Ascorbic Acid.

\pagebreak  

**- Means of Orange Juice by dose**  

```{r,echo=FALSE}
Global_OJ

```  

**- Means of Ascorbic Acid by dose**  

```{r,echo=FALSE}
Global_VC
```  

However, note that at 2 mg/day, there is no difference in the tooth growth between OJ or VC absorption (26.06 VS 26.14).  

## 4. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose  

**- Hypothesis 1 : mean Orange juice = mean absorbic acid in the tooth growth**  

```{r}
t.test(len~supp, paired = FALSE, var.equal = FALSE, mdata)
```  

With a p-value equals to 6% slightly above than the threshold of 5% and confidence intervals which include 0 then the null hypothesis cannot be rejected.  

As above we can not reject the NULL hypothesis, we go in further details by dose if now we can reject the NULL hypothesis.  

\pagebreak  

**- Hypothesis 1.1 : mean Orange juice = mean absorbic acid in the tooth growth with dose = 0.5 mg/day**  

```{r}
t.test(len~supp, paired = FALSE, var.equal = FALSE, subset(mdata, dose == 0.5))
```  

With a p-value equals to 0.6% well below than the threshold of 5% and confidence intervals which not include 0 then the null hypothesis can be rejected. Thus orange juice mean is above of the absorbic mean with a daily dose of 0.5 mg/day (as we expected).  

**- Hypothesis 1.2 : mean Orange juice = mean absorbic acid in the tooth growth with dose = 1 mg/day**  

```{r}
t.test(len~supp, paired = FALSE, var.equal = FALSE, subset(mdata, dose == 1))
```  

With a p-value equals to 0.1% well below than the threshold of 5% and confidence intervals which not include 0 then the null hypothesis can be rejected. Thus orange juice mean is above of the absorbic mean with a daily dose of 1 mg/day (as we expected).  

**- Hypothesis 1.3 : mean Orange juice = mean absorbic acid in the tooth growth with dose = 2 mg/day**  

```{r}
t.test(len~supp, paired = FALSE, var.equal = FALSE, subset(mdata, dose == 2))
```  

With a p-value equals to 96% well above than the threshold of 5% and confidence intervals which include 0 then the null hypothesis cannot be rejected.  

**- Hypothesis 1.4 : mean Orange juice = mean absorbic acid in the tooth growth without dose = 2 mg/day**  


```{r}
t.test(len~supp, paired = FALSE, var.equal = FALSE, subset(mdata, dose != 2))
```  

With a p-value equals to 0,4% well below than the threshold of 5% and confidence intervals which not include 0 then the null hypothesis can be rejected.  
  
With this hypothesis we can understand why the null hypothesis 1 can not be rejected, the fault of the dose of 2 mg/day in the data set.  

---------   

Now we are going to verify like the previous analysis, if our plot analyzes are corrects. Here we want to compare the mean by dose without take in account the delivery methods.  

First of all I want to define the correlation function between *dose* and *len*.  

```{r}
cor(mdata$dose,mdata$len)
```  
  
The two variables have 80% of correlation which is well positive. Even if 60 observations in the sample is correct to conclude of the relationship between the variables, it's not enough and need to be verify with a test hypothesis.

\pagebreak  

**- Hypothesis 2.1 : mean 0.5 mg/day = mean 1 mg/day in the tooth growth**  

```{r}
dose1 <- subset(mdata, dose %in% c(0.5,1))
t.test(len~dose, paired = FALSE, var.equal = FALSE, dose1)
```  

With a p-value equals to 1.268e-07 widely well below of 0.5 and a confidence interval without include 0, the NULL hypothesis that the mean of dose 0.5 and 1 mg/day are equals can be rejected. Mean of 1 mg/day is well above of mean of 0.5 mg/day.  

**- Hypothesis 2.2 : mean 1 mg/day = mean 2 mg/day in the tooth growth**  

As we are certificate over 0.5 vs 1 mg/day, we are going to do it over 1 vs 2 mg/day. It's not necessary to do 0.5 vs 2 mg/day as the hypothesis 2.1 have already give us an analysis until 1 mg/day.

```{r}
dose2 <- subset(mdata, dose %in% c(1,2))
t.test(len~dose, paired = FALSE, var.equal = FALSE, dose2)
```  

As we expected, the p-value is very low, well well below of 5% and there is no 0 in the confidence interval, thus the NULL hypothesis is rejected.  

## Conclusion  

At the first step we have used the exploratory analysis thanks to plots to understand the data set and the variables attached. The first plot shows us an effect of the tooth growth between the dose levels of vitamin C (0.5, 1, 2 mg/day). More the level is important, more the tooth grow. The second plot shows us the impact of the delivery methods of the vitamin C on the tooth growth. We have conclude that Orange Juice gives better results than Ascorbic Acid method delivery. Even if we have seen a decrease impact over the dose level on the Orange Juice because at 2 mg/day, VC is going back at the same mean lenght of tooth than OJ. It could be interesting to add 2 more level doses above 2 mg/day to see if VC get the upper hand over OJ.
But exploratory analysis need to be support with statistical analysis so I made it on the same area and its leads us to conclude that the dose of vitamin C and the delivery methods have clearly an impact over the tooth growth.